---
title: "The tidycat package: expand broom::tidy() output for categorical parameter estimates"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The tidycat package: expand broom::tidy() outputs for categorical parameter estimates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

The `tidycat` package includes the `tidy_categorical()` function to expand `broom::tidy()` outputs for categorical parameter estimates.

## Additional columns on broom::tidy() for categorical parameter estimates

The `tidy()` function in the broom package takes the messy output of built-in functions in R, such as `lm()`, and turns them into tidy data frames.

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(broom)
m0 <- esoph %>%
   mutate_if(is.factor, ~factor(., ordered = FALSE)) %>%
   glm(cbind(ncases, ncontrols) ~ agegp + tobgp + alcgp, data = ., family = binomial())
# tidy
tidy(m0)
```
Note: Currently ordered factor not supported in `tidycat`, hence their removal in `mutate_if()` above

The `tidy_categorical()` function adds further columns (`variable`, `level`, `effect` and `reference`) to the `broom::tidy()` output to help manage categorical variables

```{r}
library(tidycat)
m0 %>%
  tidy() %>%
  tidy_categorical(m = m0, include_reference =  FALSE)
```

Include additional rows for reference category terms and column to indicate their location by setting `include_reference = TRUE` (default). Setting `exponentiate = TRUE` ensures the parameter estimates in the reference group are set to one instead of zero (even odds in the logistic regression example below).

```{r}
m0 %>%
  tidy(exponentiate = TRUE) %>%
  tidy_categorical(m = m0, exponentiate = TRUE, reference_label = "Baseline") %>%
  select(-statistic, -p.value)
```

## Standard coefficient plots

The results from `broom::tidy()` can be used to quickly plot estimated coefficients and their confidence intervals.

```{r, fig.width=6, fig.height=4}
# store parameter estimates and confidence intervals (except for the intercept)
d0 <- m0 %>%
  tidy(conf.int = TRUE) %>%
  slice(-1)
d0

library(ggplot2)
library(tidyr)
ggplot(data = d0,
        mapping = aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
   coord_flip() +
   geom_hline(yintercept = 0, linetype = "dashed") +
   geom_pointrange()
```

## Enhanced coefficient plots

Using the additional columns from `tidy_categroical()` the coefficient plot can be enhanced to include the reference category, allowing the reader to better grasp the meaning of the parameter estimates in each categorical variable. Using `ggforce::facet_col()` the terms of each variable can be separated to further improve the presentation of the coefficient plot. 

```{r, fig.width=6, fig.height=4}
d0 <- m0 %>%
  tidy(conf.int = TRUE) %>%
  tidy_categorical(m = m0) %>%
  slice(-1)

d0 %>%
  select(-(3:5))

library(ggforce)
ggplot(data = d0,
        mapping = aes(x = level, y = estimate, colour = reference,
                      ymin = conf.low, ymax = conf.high)) +
   facet_col(facets = vars(variable), scales = "free_y", space = "free") +
   coord_flip() +
   geom_hline(yintercept = 0, linetype = "dashed") +
   geom_pointrange()
```

Note the switch of the `x` aesthetic to the `level` column rather than `term`.

Alternatively, horizontal plots can be obtained using `ggforce::facet_row()` and loosing `coord_flip()`;

```{r, fig.width=6, fig.height=4}
ggplot(data = d0,
      mapping = aes(x = level, y = estimate,
                    ymin = conf.low, ymax = conf.high,
                    colour = reference)) +
 facet_row(facets = vars(variable), scales = "free_x", space = "free") +
 geom_hline(yintercept = 0, linetype = "dashed") +
 geom_pointrange() +
 theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


## Interactions

Models with interactions can also be handled in `tidy_categorical()`. Using the `mtcars` data we can create three types of interactions (between two numeric variables, between a numeric variable and categorical variable and between two categorical variables)

```{r}
m1 <- mtcars %>%
  mutate(engine = recode_factor(vs, `0` = "straight", `1` = "V-shaped"),
         transmission = recode_factor(am, `0` = "automatic", `1` = "manual")) %>%
  lm(mpg ~ as.factor(cyl) + wt * hp + wt * transmission + engine * transmission , data = .)

tidy(m1)
```
Setting `n_level = TRUE` creates an additional column to monitor the number of observations in each of level of the categorical variables, including interaction terms in the model:

```{r}
d1 <- m1 %>%
  tidy(conf.int = TRUE) %>%
  tidy_categorical(m = m1, n_level = TRUE) %>%
  slice(-1)

d1 %>%
  select(-(2:7))
```

We can use similar plotting code as above to plot the interactions:

```{r, fig.width=6, fig.height=6}
ggplot(data = d1,
        mapping = aes(x = level, y = estimate, colour = reference,
                      ymin = conf.low, ymax = conf.high)) +
   facet_col(facets = vars(variable), scales = "free_y", space = "free") +
   coord_flip() +
   geom_hline(yintercept = 0, linetype = "dashed") +
   geom_pointrange()
```

The empty levels can be dropped by filtering on the `n_level` column for categories with more than zero observations and not `NA` in term column.

```{r, fig.width=6, fig.height=5}
ggplot(data = d1 %>% 
         filter(n_level > 0 | !is.na(term)),
        mapping = aes(x = level, y = estimate, colour = reference,
                      ymin = conf.low, ymax = conf.high)) +
   facet_col(facets = vars(variable), scales = "free_y", space = "free") +
   coord_flip() +
   geom_hline(yintercept = 0, linetype = "dashed") +
   geom_pointrange()
```
